---
title: "Hello, Dolt!"
author: "Noam Ross"
date: "9/1/2021"
output: github_document
---

```{r packages, results = FALSE, message = FALSE}
library(DBI)
library(RMariaDB)
library(duckdb)
library(sys)
library(nycflights13)
library(withr)
library(arrow)
library(dplyr)
library(dbplyr)
library(bench)
library(dbx)
library(waldo)
library(vroom)
knitr::opts_chunk$set(error = FALSE)
```

# Setting up a Dolt Database

Install dolt from source. The latest source version is needed

```{bash eval = FALSE}
git clone git@github.com:dolthub/dolt.git
cd dolt/go && go install ./cmd/dolt ./cmd/git-dolt ./cmd/git-dolt-smudge
```
Check current version (shell)

```{r, eval = TRUE}
exec_wait("~/go/bin/dolt", "version")
```

<!-- Kill stuff from previous runs -->

```{bash, eval=FALSE, include = FALSE}
kill $(lsof -t -i:3333)
```

```{r, eval=TRUE, include = FALSE}
unlink("doltdb", recursive = TRUE)
unlink("duckdb")
unlink("duckdb.wal")
unlink("doltdb", recursive = TRUE)
unlink("flights.parquet")
```

.
## Start a server and connect to it

```{r, eval = TRUE}
dir.create("doltdb", showWarnings = FALSE)
with_dir("doltdb", exec_wait("~/go/bin/dolt", "init"))
dolt_server_pid <- with_dir("doltdb", sys::exec_background(
  "~/go/bin/dolt", c("sql-server", "--port 3333", "--host 127.0.0.1",
                     "--user user", "--password pwd", "--l error")
))
Sys.sleep(1) # Wait for server to fully start up before connecting
```

```{r}
dolt_conn <- dbConnect(RMariaDB::MariaDB(), host = "127.0.0.1", port = 3333,
                       username = "user", password = "pwd", dbname = "doltdb")
```

# A Grossly Unfair Performance Comparison

Comparing Dolt to DuckDB and using arrow to write Parquet files. In our use case,
we might use the latter options and use S3 versioning to track those binary files. 

## Writing to the database

Dolt

```{r}
flights <- nycflights13::flights
dim(flights)
print(object.size(flights), units = "Mb")

# Using dbx::dbInsert here rather than dbWriteTable because it is much faster
# to generate large INSERT statements than iterating through a parameterized
# statement as dbBind does.
dbCreateTable(dolt_conn, "flights", flights)
dolt_write_time <- system.time(
  dbxInsert(dolt_conn, "flights", flights)
)
dolt_write_time
```

DuckDB

```{r}
duck_conn <- dbConnect(duckdb::duckdb(), dbdir = "duckdb")
duck_write_time <- system.time(
  dbWriteTable(duck_conn, "flights", flights, overwrite = TRUE)
)
duck_write_time
```

Parquet

```{r}
pq_write_time <- system.time(
  arrow::write_parquet(flights, "flights.parquet")
)
pq_write_time
```

This isn't really a "fair" comparison, given the different goals of
the projects, but the write time for Dolt is `r signif(dolt_write_time[3] / duck_write_time[3], 1)`
time that of DuckDB, and `r signif(dolt_write_time[3] / pq_write_time[3], 1)` of
writing to Parquet via arrow.  

## Read comparisons and dbplyr testing

```{r}
dolt_read_time <- system.time(
  fl_dolt <- tbl(dolt_conn, "flights") |> 
    collect()
)
dolt_read_time
```

```{r}
duck_read_time <- system.time(
  fl_duck <- tbl(duck_conn, "flights") |> 
    collect()
)
duck_read_time
```

```{r}
pq_read_time <- system.time(
  fl_parquet  <- read_parquet("flights.parquet")
)
pq_read_time
```

For reading the whole table, Dolt read time is `r signif(dolt_read_time[3] / duck_read_time[3], 1)`
times that of DuckDB, and `r signif(dolt_write_time[3] / pq_read_time[3], 1)` times
that of reading parquet.

Now let's compare querying and reading a tiny value

```{r}
query_benchmark <- bench::mark(
  dolt_query = tbl(dolt_conn, "flights") |> 
    filter(carrier == "MQ", dest == "DTW", month == 5, arr_delay < -10, day == 30 ) |> 
    select(flight, tailnum) |> 
    collect(),
  
  duck_query = tbl(duck_conn, "flights") |> 
    filter(carrier == "MQ", dest == "DTW", month == 5, arr_delay < -10, day == 30 ) |> 
    select(flight, tailnum) |> 
    collect(),
  
  pq_query = arrow::open_dataset("flights.parquet") |> 
    filter(carrier == "MQ", dest == "DTW", month == 5, arr_delay < -10, day == 30 ) |> 
    select(flight, tailnum) |> 
    collect(),
  min_iterations = 10,
  check = FALSE # data are returned in different order, so don't check they are identical
)
query_benchmark
```

<!-- Wrap up -->
```{r, include = FALSE}
dbDisconnect(dolt_conn)
tools::pskill(dolt_server_pid)
try(duckdb_shutdown(duckdb(dbdir = "duckdb")), silent = TRUE)
unlink("duckdb")
unlink("duckdb.wal")
unlink("doltdb")
unlink("flights.parquet")
```

